Hierarchically Supervised Latent Dirichlet Allocation

### 2011-06-02 paper/abstract.tex
misc

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
model for hierarchically and multiply labeled bag-of-word data.  Examples of
such data include web pages and their placement in directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data is also of interest.
%We define a model that uses probit regressors on a conditionally dependent
%label hierarchy tied to latent Dirichlet allocation (LDA). 
We demonstrate HSLDA on large-scale data from clinical document labeling and
retail product categorization tasks. We show that leveraging the structure from
hierarchical labels improves out-of-sample label prediction substantially when
compared to models that do not. %Furthermore, we show evidence
%that the resulting HSLDA topics are more descriptive of the underlying data
%than sLDA topics, which ignore the label hierarchy.

>>>>

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
model for hierarchically and multiply labeled bag-of-word data.  Examples of
such data include web pages and their placement in directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data are also of interest.
%We define a model that uses probit regressors on a conditionally dependent
%label hierarchy tied to latent Dirichlet allocation (LDA). 
We demonstrate HSLDA on large-scale data from clinical document labeling and
retail product categorization tasks. We show that leveraging the structure from
hierarchical labels improves out-of-sample label prediction substantially when
compared to models that do not. %Furthermore, we show evidence
%that the resulting HSLDA topics are more descriptive of the underlying data
%than sLDA topics, which ignore the label hierarchy.

### 2011-06-02 paper/introduction.tex
misc

There exist many sources of unstructured data that have been partially or
completely categorized by human editors.  In this paper we focus on
unstructured textual data that has been, at least in part, manually
categorized.  Examples include but are not limited to webpages and curated
hierarchical directories of the same \citep{DMOZ}, product descriptions and
catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes
(e.g.~hospital discharge summaries with 
International Classification of Disease 9th Revision, Clinical Modification
(ICD-9-CM) codes assigned \cite{ICD9}).  In this work we show how to combine
these two sources of information using a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. The models and techniques that we develop in
this paper are applicable in other domains as well, namely, any unstructured
representations of data that have been hierarchically classified (e.g.,~image
catalogs with bag-of-feature image representations).

@@

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. For the sake of
simplicity, we focus on ISA hierarchies, but the model can be
applied to other hierarchies. We extend supervised latent Dirichlet
allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
supervision. We propose an efficient way to incorporate hierarchical
information into the model. We hypothesize that the context of labels within
the hierarchy provides valuable information about labeling.

>>>>

There exist many sources of unstructured data that have been partially or
completely categorized by human editors.  In this paper we focus on
unstructured text data that has been, at least in part, manually
categorized.  Examples include but are not limited to webpages and curated
hierarchical directories of the same \citep{DMOZ}, product descriptions and
catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
and patient records %hospital treatment transcripts
and diagnosis codes assigned to them for bookkeeping and insurance purposes
(e.g.~hospital discharge summaries with 
International Classification of Disease 9th Revision, Clinical Modification
(ICD-9-CM) codes assigned \cite{ICD9}).  In this work we show how to combine
these two sources of information using a single model that allows one to
categorize new text documents automatically, suggest labels that might be
inaccurate, compute improved similarities between documents for information
retrieval purposes, and more. The models and techniques that we develop in
this paper are applicable in other data as well, namely, any unstructured
representations of data that have been hierarchically classified (e.g.,~image
catalogs with bag-of-feature image representations).

@@

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. For the sake of
simplicity, we focus on is-a hierarchies, but the model can be
applied to other hierarchies. We extend supervised latent Dirichlet
allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
supervision. We propose an efficient way to incorporate hierarchical
information into the model. We hypothesize that the context of labels within
the hierarchy provides valuable information about labeling.

### 2011-06-02 paper/model.tex
Worked references and cleaned them up significantly.
misc

HSLDA is a model for hierarchically, multiply-labeled, bag-of-words data.  We will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d} \in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document $d$.  Let there be $D$ such documents and let the size of the vocabulary be $V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label except root labels $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
 We will for exposition purposes assume that this label set has hard ``is-a'' parent-child constraints (explained later), although this assumption can be relaxed at the cost of more complicated inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we consider, only positive label applications are observed.

@@

The second half of step 4 is a substantial part of our contribution to the general class of supervised LDA models.  Here each document is generatively labeled using a hierarchy of conditionally dependent probit regressors \cite{gelman?,probit_regression?}.  For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.  Note that label $y_{l,d}$ can only be applied to document $d$ if its parent label $\mathrm{pa}(l)$ is also applied (these expressions are specific to is-a constraints but can be modified to accommodate different constraints).  The regression coefficients $\boldsymbol\eta_l$ are independent a priori, however, the hierarchical coupling in this model induces a posteriori dependence.   The net effect of this is that label predictors deeper in the label hierarchy are able to focus on finding specific, conditional labeling features.  We believe this to be a significant source of the empirical label prediction improvement we observe experimentally.  We test this hypothesis in Section~\ref{sec:experiments}.

>>>>

HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.  We will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d} \in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document $d$.  Let there be $D$ such documents and let the size of the vocabulary be $V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label except root labels $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
 We will for exposition purposes assume that this label set has hard ``is-a''
 parent-child constraints (explained later), although this assumption can be
 relaxed at the cost of more computationally complex inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we consider, only positive label applications are observed.

@@

The second half of step 4 is a substantial part of our contribution to the general class of supervised LDA models.  Here each document is generatively labeled using a hierarchy of conditionally dependent probit regressors \cite{gelmanbda04}.  For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.  Note that label $y_{l,d}$ can only be applied to document $d$ if its parent label $\mathrm{pa}(l)$ is also applied (these expressions are specific to is-a constraints but can be modified to accommodate different constraints).  The regression coefficients $\boldsymbol\eta_l$ are independent a priori, however, the hierarchical coupling in this model induces a posteriori dependence.   The net effect of this is that label predictors deeper in the label hierarchy are able to focus on finding specific, conditional labeling features.  We believe this to be a significant source of the empirical label prediction improvement we observe experimentally.  We test this hypothesis in Section~\ref{sec:experiments}.
